---
title: "Create Powerful Model Explanations with Logistic Regression"
author: "Jin Cui"
date: "2023-01-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import Libraries
```{r Import library}
library(readxl)
library(data.table)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(writexl)
library(ggpubr)
library(scales)
library(plotly)
library(officer)
library(rvg)

## for model fitting
library(corrplot)
library(DALEX)
library(car)
library(rpart)
library(rpart.plot)
library(pROC)
library(pdp)
library(caret)
library(randomForest)
library(mlr)
#library(xgboost)
#library(SHAPforxgboost)

```

#     1 - Data Import and EDA

## Set up and ingest data
```{r set up data paths}

data_raw <- read.csv('C:/Users/Jin/Documents/00_Machine_Learning/Logistic_Regression/Data/WA_Fn-UseC_-Telco-Customer-Churn.csv', header = TRUE)

#str(data_raw)


```

## Map categorical columns
```{r Data Transformation - Create Categorical Columns}

setDT(data_raw)

data_raw[, SeniorCitizen_Flag := case_when(
                                 `SeniorCitizen` == 0 ~ "No",
                                 `SeniorCitizen` == 1 ~ "Yes" 
               
                                 )]


data_raw[, Churn_Flag := case_when(
                                 `Churn` == "Yes" ~ TRUE,
                                 `Churn` == "No" ~ FALSE 
               
                                 )]

data_raw$Bundle_Flag <- as.factor(ifelse (data_raw$PhoneService == "Yes" & data_raw$InternetService != "No" ,"Yes","No"))

```


## Apply filters
```{r Apply filters for Inforce vs. Churn}

data_if    <- data_raw %>%
  filter(`Churn` == "No")

data_churn <- data_raw %>%
  filter(`Churn` == "Yes")

```






## Density Plots for Continuous Variables



### Monthly Charge
```{r Density Plot for Monthly Charges}

limit_prem <- 120

den_prem <- ggplot() + 
  geom_density(data = data_if, aes(x = data_if[["MonthlyCharges"]], y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6, color = NA) +
  geom_density(data = data_churn, aes(x = data_churn[["MonthlyCharges"]], y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6, color = NA) +
  
#scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
scale_x_continuous(labels = comma, breaks = seq(0, limit_prem, 10) 
  #                   , guide = guide_axis(n.dodge = 2)
                   ) +
scale_y_continuous(labels = percent, breaks=seq(0, 0.006, 0.001)) +

  theme(
        axis.title = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 8)
        ) +
  labs(
       #x = "Full Lapse Indicator", y = "Exposure %", 
       title = "Distribution of Churn by Monthly Charges") +
  #guides(fill = guide_legend("Full Lapse Indicator")) +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        #panel.grid.major.y = element_line(color = 'black', linetype = 'dotted'),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank() )


box_prem <- ggplot(data_raw, aes(x = as.factor(data_raw[["Churn"]]), y = data_raw[["MonthlyCharges"]], fill = as.factor(data_raw[["Churn"]]))) + 
  geom_boxplot(outlier.colour=NA, alpha = 0.6) +
  scale_y_continuous(labels = comma, breaks=seq(0, limit_prem, 10)
                     #, guide = guide_axis(n.dodge = 2)
                     ) +
  coord_cartesian(ylim = c(0, 200)) +
  theme(legend.title = element_blank(),
        legend.position = "none"
        ) +
  labs(x = "Churn") +
        #x = "Full Lapse Indicator", y = "API", 
        #title = "Distribution of Lapse by Premium") + 
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8)
        ) +
  scale_fill_manual(values=c("#AAD8C8", "#2FB3B6")) +
  stat_summary(fun = "mean", shape = 4, color = "#303E46")
  


ggr_prem_den <- ggarrange(den_prem, box_prem, widths = c(2.0, 0.7),
          #labels = c("A", "B", "C"),
          ncol = 2, nrow = 1)

#dev.new()
print(ggr_prem_den)

```


### Tenure (Months)
```{r Density Plot for Tenure}

limit_tenure <- 75

den_tenure <- ggplot() + 
  geom_density(data = data_if, aes(x = data_if[["tenure"]], y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6, color = NA) +
  geom_density(data = data_churn, aes(x = data_churn[["tenure"]], y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6, color = NA) +
  
#scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
scale_x_continuous(labels = comma, breaks = seq(0, limit_tenure, 5) 
  #                   , guide = guide_axis(n.dodge = 2)
                   ) +
scale_y_continuous(labels = percent, breaks=seq(0, 0.006, 0.001)) +

  theme(
        axis.title = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 8)
        ) +
  labs(
       #x = "Full Lapse Indicator", y = "Exposure %", 
       title = "Distribution of Churn by Tenure (Months)") +
  #guides(fill = guide_legend("Full Lapse Indicator")) +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        #panel.grid.major.y = element_line(color = 'black', linetype = 'dotted'),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank() )


box_tenure <- ggplot(data_raw, aes(x = as.factor(data_raw[["Churn"]]), y = data_raw[["tenure"]], fill = as.factor(data_raw[["Churn"]]))) + 
  geom_boxplot(outlier.colour=NA, alpha = 0.6) +
  scale_y_continuous(labels = comma, breaks=seq(0, limit_tenure, 5)
                     #, guide = guide_axis(n.dodge = 2)
                     ) +
  coord_cartesian(ylim = c(0, 200)) +
  theme(legend.title = element_blank(),
        legend.position = "none"
        ) +
  labs(x = "Churn") +
        #x = "Full Lapse Indicator", y = "API", 
        #title = "Distribution of Lapse by Premium") + 
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8)
        ) +
  scale_fill_manual(values=c("#AAD8C8", "#2FB3B6")) +
  stat_summary(fun = "mean", shape = 4, color = "#303E46")
  


ggr_tenure_den <- ggarrange(den_tenure, box_tenure, widths = c(2.0, 0.7),
          #labels = c("A", "B", "C"),
          ncol = 2, nrow = 1)

#dev.new()
print(ggr_tenure_den)

```


## Histogram Plots for Categorical Variable (on same scale)


### Gender
```{r }


labels_Gender <- c("Female", "Male")

Gender_hist <- ggplot() + 
  geom_bar(data = data_if, aes(x = as.factor(data_if[["gender"]]), y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6) +
    geom_bar(data = data_churn, aes(x = as.factor(data_churn[["gender"]]), y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6) +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.1)) +
  scale_x_discrete(labels = labels_Gender) +
  #scale_fill_brewer(palette = "Pastel1") +
  #scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
  theme(axis.title = element_blank()) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  labs(title = "Churn Distribution by Gender") +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 9))

print(Gender_hist)

```



### Senior Citizen Flag
```{r }


labels_SC <- c("Yes", "No")

SC_hist <- ggplot() + 
  geom_bar(data = data_if, aes(x = as.factor(data_if[["SeniorCitizen_Flag"]]), y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6) +
    geom_bar(data = data_churn, aes(x = as.factor(data_churn[["SeniorCitizen_Flag"]]), y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6) +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.1)) +
  scale_x_discrete(labels = labels_SC) +
  #scale_fill_brewer(palette = "Pastel1") +
  #scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
  theme(axis.title = element_blank()) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  labs(title = "Churn Distribution by Senior Citizen Flag") +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 9))

print(SC_hist)

```


### Phone Service
```{r }


labels_PS <- c("Yes", "No")

PS_hist <- ggplot() + 
  geom_bar(data = data_if, aes(x = as.factor(data_if[["PhoneService"]]), y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6) +
    geom_bar(data = data_churn, aes(x = as.factor(data_churn[["PhoneService"]]), y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6) +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.1)) +
  scale_x_discrete(labels = labels_PS) +
  #scale_fill_brewer(palette = "Pastel1") +
  #scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
  theme(axis.title = element_blank()) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  labs(title = "Churn Distribution by Phone Service Flag") +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 9))

print(PS_hist)

```


### Internet Service
```{r }


labels_IS <- c("DSL", "Fiber optic", "No")

IS_hist <- ggplot() + 
  geom_bar(data = data_if, aes(x = as.factor(data_if[["InternetService"]]), y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6) +
    geom_bar(data = data_churn, aes(x = as.factor(data_churn[["InternetService"]]), y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6) +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.1)) +
  scale_x_discrete(labels = labels_IS) +
  #scale_fill_brewer(palette = "Pastel1") +
  #scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
  theme(axis.title = element_blank()) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  labs(title = "Churn Distribution by Internet Service Flag") +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 9))

print(IS_hist)

```



### Contract
```{r }


labels_Contract <- c("Monthly", "One Year", "Two Year")

Contract_hist <- ggplot() + 
  geom_bar(data = data_if, aes(x = as.factor(data_if[["Contract"]]), y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6) +
    geom_bar(data = data_churn, aes(x = as.factor(data_churn[["Contract"]]), y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6) +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.1)) +
  scale_x_discrete(labels = labels_Contract) +
  #scale_fill_brewer(palette = "Pastel1") +
  #scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
  theme(axis.title = element_blank()) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  labs(title = "Churn Distribution by Contract Type") +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 9))

print(Contract_hist)

```



### Paperless Billing
```{r }


labels_PB <- c("Yes", "No")

PB_hist <- ggplot() + 
  geom_bar(data = data_if, aes(x = as.factor(data_if[["PaperlessBilling"]]), y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6) +
    geom_bar(data = data_churn, aes(x = as.factor(data_churn[["PaperlessBilling"]]), y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6) +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.1)) +
  scale_x_discrete(labels = labels_PB) +
  #scale_fill_brewer(palette = "Pastel1") +
  #scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
  theme(axis.title = element_blank()) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  labs(title = "Churn Distribution by Paperless Billing Flag") +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 9))

print(PB_hist)

```


### Payment Method
```{r }


labels_PM <- c("Bank Transfer", "Credit Card", "Electronic Cheque", "Mailed Cheque")

PM_hist <- ggplot() + 
  geom_bar(data = data_if, aes(x = as.factor(data_if[["PaymentMethod"]]), y = ..count../sum(..count..)), fill = "#AAD8C8", alpha = 0.6) +
    geom_bar(data = data_churn, aes(x = as.factor(data_churn[["PaymentMethod"]]), y = ..count../sum(..count..)), fill = "#2FB3B6", alpha = 0.6) +
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 1, 0.1)) +
  scale_x_discrete(labels = labels_PM, guide = guide_axis(n.dodge = 2)) +
  #scale_fill_brewer(palette = "Pastel1") +
  #scale_fill_manual(values = c("#AAD8C8", "#2FB3B6")) +
  theme(axis.title = element_blank()) +
  theme(legend.title = element_blank(),
        legend.position = "none") +
  labs(title = "Churn Distribution by Payment Method") +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 9))

print(PM_hist)

```


## Stat2D plot


### Monthly Charges by Contract Type STAT2D plot
```{r }

labels_y_contract_type <- c("Month-to-Month", 
                           "One Year", 
                           "Two Year")

Stat2D_MC_contract <- ggplot(data_if, aes(x = `MonthlyCharges`, y = `Contract`)) +

  stat_bin2d(bins = 20) +
  scale_fill_gradient(low = "#EFEFD2", high = "#F15A24", limits = c(0, 1000)) +
 scale_x_continuous(labels = comma, breaks=seq(0, 120, 10)
                     ,guide = guide_axis(n.dodge = 2)
                     ) +
 scale_y_discrete(labels = labels_y_contract_type) +

  theme(axis.title = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 8)
        ) +
  labs(title = "Distribution of Contract Type by Monthly Charges") +
  #guides(fill = guide_legend(title = "Full Lapse")) +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        #panel.grid.major.y = element_line(color = 'black', linetype = 'dotted'),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        #axis.ticks = element_blank() 
        ) 

print(Stat2D_MC_contract)

```

### Monthly Charges by Payment Method STAT2D plot

```{r }

labels_y_payment_method <- c("Bank Transfer", 
                           "Credit Card", 
                           "Electronic Cheque",
                           "Mailed Cheque")

Stat2D_MC_PM <- ggplot(data_if, aes(x = `MonthlyCharges`, y = `PaymentMethod`)) +

  stat_bin2d(bins = 20) +
  scale_fill_gradient(low = "#EFEFD2", high = "#F15A24", limits = c(0, 500)) +
 scale_x_continuous(labels = comma, breaks=seq(0, 120, 10)
                     ,guide = guide_axis(n.dodge = 2)
                     ) +
 scale_y_discrete(labels = labels_y_payment_method) +

  theme(axis.title = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 8)
        ) +
  labs(title = "Distribution of Payment Method by Monthly Charges") +
  #guides(fill = guide_legend(title = "Full Lapse")) +
  theme(panel.background = element_rect(fill = '#FFFFFF', color = "#FFFFFF"),
        panel.grid.major.x = element_blank(),
        #panel.grid.major.y = element_line(color = 'black', linetype = 'dotted'),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        #axis.ticks = element_blank() 
        ) 

print(Stat2D_MC_PM)

```


#     2 - GLM Fit


### Train Test Split

```{r}

y = data_raw$Churn_Flag


set.seed(268)

sample_size <- floor(0.7 * nrow(data_raw))
sample_indi <- sample(seq_len(nrow(data_raw)), size = sample_size)

d_train <- data_raw[sample_indi,]
d_test <- data_raw[-sample_indi,]
y_train <- y[sample_indi]
y_test <- y[-sample_indi]


```

### GLM Fit 1 - Base
```{r}


glm_1 <- glm(Churn_Flag ~
                  
  tenure
+ MonthlyCharges

+ relevel(factor(gender), ref = "Female")
+ relevel(factor(SeniorCitizen_Flag), ref = "Yes")
+ relevel(factor(PhoneService), ref = "No")
+ relevel(factor(InternetService), ref = "DSL")
+ relevel(factor(Contract), ref = "Month-to-month")
+ relevel(factor(PaperlessBilling), ref = "No")
+ relevel(factor(PaymentMethod), ref = "Bank transfer (automatic)")

#+ Partner
#+ Dependents
#+ MultipleLines
#+ OnlineSecurity	
#+ OnlineBackup 
#+ DeviceProtection
#+	TechSupport
#+	StreamingTV
#+	StreamingMovies




, data = d_train
, family = binomial("logit")

)


summary(glm_1)
#write.csv(summary(glm_1)$coefficients, "summary_glm_1_telco_churn.csv")

```

### Model Diagnostics

```{r}

p_train_glm_1 <- predict(glm_1, d_train, type = "response")
p_test_glm_1  <- predict(glm_1, d_test, type = "response")

mean(y_train)
mean(p_train_glm_1)

mean(y_test)
mean(p_test_glm_1)


thresh <- quantile(p_train_glm_1, 1 - mean(y_train))
#thresh_LS

confusionMatrix(factor(p_train_glm_1 > thresh), factor(y_train), positive = "TRUE")
confusionMatrix(factor(p_test_glm_1 > thresh), factor(y_test), positive = "TRUE")


roc(y_train, p_train_glm_1, plot = T, ci = T)
roc(y_test, p_test_glm_1, plot = T, ci = T)

```

